---
title: "Flight delays"
output: html_document
runtime: shiny
---
  
```{r echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(flights2)
```

```{r echo=FALSE}
selectInput("origin", "Select Origin:", 
            choices = sort(unique(flights2$origin)),
            selected = "JFK")

selectInput("dest", "Select Destination:",
            choices = sort(unique(flights2$dest)),
            selected = "BOS")
```

```{r echo=FALSE, message=FALSE}
trip <- reactive({
  flights2 %>%
    filter(origin == input$origin, dest == input$dest)
})

hour_avgs <- reactive({
  trip() %>% 
    group_by(hour) %>% 
    summarise(avg_delay = round(mean(arr_delay), 2))
})

hour <- reactive({
  smallest <- hour_avgs()$hour[which.min(hour_avgs()$avg_delay)]
  paste0(smallest %% 12, ifelse(smallest %/% 12, ":00 pm", ":00 am"))
})
```


# Goal

What is the best time to fly from `r renderText({input$origin})` airport to `r renderText({input$dest})`?

# Motivation

Flight delays are one of the most frustrating aspects of airline travel (well, one of many). Travellers can do little to avoid flight delays; travellers usually cannot choose where they need to travel to, or where they need to travel from. However, travelers can choose _when_ they travel.

This study will examine the correlation between _departure time_, the time a flight is scheduled to depart, and _arrival delay_, the amount of time that passes between when a plane is scheduled to arrive and when it actually arrives. Travelers can use this information to mitigate the expected delay associated with their flight.

# Data

We will examine all flights that originated from a New York City airport in 2013. The data is provided by the Bureau of Transportation Statistics (BTS) and is available for download from <http://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=236>.

# Results

The table below lists the mean arrival delay by hour.

```{r echo=FALSE}
renderTable({
  hour_avgs()
})
```

The expected delay fluctuates throughout the day. Here the delay is described by a smooth curve fitted to the data.

```{r echo=FALSE, message=FALSE, warning=FALSE}
renderPlot({
  qplot(clock, arr_delay, data = trip(), geom = "smooth") +
    xlab("Scheduled departure time") +
    ylab("Expected arrival delay (m)")
})
```

# Conclusion

The best time to fly from `r renderText({input$origin})` airport to `r renderText({input$dest})` is the hour immediately after **`r renderText({hour()})`**.

# A. Raw Data

```{r echo=FALSE, message = FALSE}
library(DT)
renderDataTable({trip()})
```

